Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach

### 2021-09-20 Documents/Bookdown/02-ext.Rmd
finish first part BF ext

The encompassing model is estimated using 6 independent chains with 10,000 iterations (warm-up 2,000). To do that we use the `brm()` fucntion from the `brms` R-package [@burknerBrmsPackageBayesian2017; burknerAdvancedBayesianMultilevel2018a], which is based on STAN [@standevelopmentteamRStanInterfaceStan2020].

@@

We can evaluate the consequences of different values choice considering the resulting prior predictions. To facilitate this step, we compute prior prediction considering only the intercept and a single parameter of interest. Remembering that the inverse link function (i.e., function that in a GLM transform the model linear prediction into the value on the original response scale) is the exponential function, we consider as intercept the value 1, because $exp(1) \approx 2.7$ that is close to the externalizing problems sample mean  `r ext_mean()`. In Table~\@ref(table-prior-predict), summary information about prior predictions for different standard deviation values are reported.
```{r table-prior-predict}
get_table_prior_predict(format = params$format)
```

@@

#### Hypotheses Matrices {-}

Finally, note that in this case, we do not need to standardize our parameters as they represent mean groups differences (see main article [TODO: add link article]).

>>>>

@@

We can evaluate the consequences of different values choice considering the resulting prior predictions. To facilitate this step, we compute prior prediction considering only the intercept and a single parameter of interest. Remembering that the inverse link function (i.e., function that in a GLM transform the model linear prediction into the value on the original response scale) is the exponential function, we consider as intercept the value 1, because $exp(1) \approx 2.7$ that is close to the externalizing problems sample mean  `r ext_mean()`. In Table~\@ref(tab:table-prior-predict), summary information about prior predictions for different standard deviation values are reported.
```{r table-prior-predict}
get_table_prior_predict(format = params$format)
```

@@

#### Posterior{-}
The encompassing model is estimated using 6 independent chains with 10,000 iterations (warm-up 2,000). To do that we use the `brm()` function from the `brms` R-package [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018a], which is based on STAN [@standevelopmentteamRStanInterfaceStan2020]. Summary of the encompassing model is presented below.
```{r}
summary(encompassing_model_ext)
```

### Hypothesis Matrices

For each informative hypothesis, we obtain an hypothesis matrix that translate equality and inequality constraints according to the encompassing model parametrization. Formalization of informative hypotheses and the procedure to derive hypothesis matrices are described in the main paper ([TODO: add link]).

Here we present the obtained hypothesis matrices were on the columns we have the parameters of the encompassing model (excluding the intercept) and each row express an equality constraint or an inequality constraint. Matrices row names follow this notation: names without square brackets indicate a constraint directly on the model parameter; names within square brackets indicate a group condition (that could be the resulting composition of more parameters). For example, `"M_Avoidant:F_Anxious"` indicates the actual interaction term of the model, whereas `"[M_Avoidant_F_Anxious]"` indicate the group condition. Equality and inequality constraints ar presented separately.

#### Null Hypothesis {-}

- **Equality matrix**, $R_{iE} = 0$ (each row is set to zero).
```{r cache=TRUE}
get_hypothesis_matrix("null", encompassing_model_ext)$eq
```

- **Inequality matrix**, $R_{iI} > 0$ (each row is set to grater than zero). There are no inequality constraints.

#### Monotropy Hypothesis {-}

- **Equality matrix**, $R_{iE} = 0$ (each row is set to zero).
```{r cache=TRUE}
get_hypothesis_matrix("monotropy", encompassing_model_ext)$eq
```

- **Inequality matrix**, $R_{iI} > 0$ (each row is set to grater than zero).
```{r cache=TRUE}
get_hypothesis_matrix("monotropy", encompassing_model_ext)$ineq
```

#### Hierarchy Hypothesis {-}

- **Equality matrix**, $R_{iE} = 0$ (each row is set to zero).
```{r cache=TRUE}
get_hypothesis_matrix("hierarchy", encompassing_model_ext)$eq
```

- **Inequality matrix**, $R_{iI} > 0$ (each row is set to grater than zero).
```{r cache=TRUE}
get_hypothesis_matrix("hierarchy", encompassing_model_ext)$ineq
```

#### Independence Hypothesis {-}

- **Equality matrix**, $R_{iE} = 0$ (each row is set to zero).
```{r cache=TRUE}
get_hypothesis_matrix("independence", encompassing_model_ext)$eq
```

- **Inequality matrix**, $R_{iI} > 0$ (each row is set to grater than zero).
```{r cache=TRUE}
get_hypothesis_matrix("independence", encompassing_model_ext)$ineq
```

#### Integration Hypothesis {-}

- **Equality matrix**, $R_{iE} = 0$ (each row is set to zero).
```{r cache=TRUE}
get_hypothesis_matrix("integration", encompassing_model_ext)$eq
```

- **Inequality matrix**, $R_{iI} > 0$ (each row is set to grater than zero).
```{r cache=TRUE}
get_hypothesis_matrix("integration", encompassing_model_ext)$ineq
```

### Centering and Adjusting

So far we have specified the encompassing prior, obtained the model posterior distribution, and defined the hypotheses matrices. Now, we need to transform our parameters of interest and center the distribution on the constraints focal points of interest. We apply the following transformation
$$
\beta = R\theta - r
$$
but we can ignore $r$ as in all our constraints it is always a vector of zeros.

Next, we get the adjusted prior and the posterior of the transformed parameters vector  $\beta$ (i.e., the parameters that identify the constraints) for each hypothesis. The adjusted prior is given by

$$
\pi_{adj}(\beta) \sim \mathcal{N}(0, \Sigma_{\beta}) = \mathcal{N}(0, R\Sigma_{\theta}R^T).
$$
Note that we set the mean vector to zero. The posterior is given by the same transformation
$$
Pr(\beta|Y) \sim \mathcal{N}(\hat{\beta}, \hat{\Sigma}_{\beta}) = \mathcal{N}(R\hat{\theta}-r, R\hat{\Sigma}_{\theta}R^T).
$$
See the main article for more details [TODO: add link].

This adjustment, however, requires the hypothesis matrix $R$ to be *full-row-rank* (i.e., all constraints are linearly independent). However, this is not the case of the hierarchy hypothesis. To overcome this issue, we follow the solution presented in the main article. First define $R^*$ selecting the maximum number of independent rows. In this case 15 contrast are independent

```{r cache=TRUE}
get_hypothesis_matrix("hierarchy", encompassing_model_ext)$hyp[1:15,]
```

```{r cache=TRUE}
R <- get_hypothesis_matrix("hierarchy", encompassing_model_ext)$hyp
```

The remaining contrasts, instead, are obtained as linear combination of the other constraints. In particular,
```{r, echo = TRUE, cache = TRUE}
# Constraint 16: [M_Anx_F_Sec - M_Sec_F_Anx]
all(R["[M_Anx_F_Sec - M_Sec_F_Anx]",] == R["[M_Anx_F_Sec]", ] - R["[M_Sec_F_Anx]", ])

# Constraint 17: [M_Av_F_Sec - M_Sec_F_Av]
all(R["[M_Av_F_Sec - M_Sec_F_Av]",] == - R["[M_Anx_F_Sec - M_Av_F_Sec]", ] + R["[M_Sec_F_Anx - M_Sec_F_Av]", ] +  R["[M_Anx_F_Sec - M_Sec_F_Anx]", ])

# Constraint 18: [M_Fear_F_Sec - M_Sec_F_Fear] 
all(R["[M_Fear_F_Sec - M_Sec_F_Fear]",] == R["[M_Av_F_Sec - M_Sec_F_Av]", ] + R["[M_Fear_F_Sec - M_Av_F_Sec]", ] - R["[M_Sec_F_Fear - M_Sec_F_Av]", ])
```

Before computing the Bayes Factor, note that we have a set of comparable hypotheses as exists a common solution to the set of linear equations obtained by setting all hypothesis constraints equal to zero. The solution is the trivial solution of simply consider all parameters equal to zero. Finally, we do not need to standardize our parameters as they represent mean groups differences. See the main article for detailed explanation [TODO: add link article].

### Results and Sensitivity

To compute the Bayes Factor we evaluate marginal densities and conditional probabilities as described in details in the main article [TODO: add link article]. Bayes Factor and posterior probability of each hypothesis are reported in Table~\@ref(tab:table-bf-results-ext).

```{r table-bf-results-ext, cache = TRUE}
get_table_bf(bf_result = BF_weights_ext,
             path_img = "images/", problem = "ext",
             format = params$format)
```

Remember, however, that prior specification affects the Bayes Factor results. Therefore, we also evaluate the results considering different prior settings. In particular, we consider as possible priors for the parameters of interest:

- $\mathcal{N}(0,.5)$ - unreasonable tight prior
- $\mathcal{N}(0,1)$ - tighter prior
- $\mathcal{N}(0,3)$ - original prior
- $\mathcal{N}(0,5)$ - more diffuse prior
- $\mathcal{N}(0,10)$ - unreasonably diffuse prior

The results of the prior sensitivity analysis are reported in Table~\@ref(tab:table-sens-prior-analysis-ext).

```{r table-sens-prior-analysis-ext, cache = TRUE}
get_table_sens_analysis(summary_sensitivity_ext,
                        format = params$format,
                        bookdown = TRUE)
```

Overall results consistently indicate the Monotropy Hypothesis as the most supported by the data. However, we can observe two distinct patterns. As the prior gets more diffuse, the order of magnitude of the Bayes Factor comparing each hypothesis with the encompassing model increases. Moreover, the probability of the Null Hypothesis increases with more diffuse prior, whereas the probabilities of the Hierarchy, Independence and Integration Hypothesis increases with tighter priors.

To interpret these patterns, remember that order constraints are insensitive to the distribution specification as long as the distribution is symmetric and centred on the constraint focal point. On the contrary, equality constraints are highly affected by the prior definition (see he main article for more details[TODO: add link article]).

All the defined hypotheses include equality constraints. Thus, for more diffuse prior we observe that the order of magnitude of the Bayes Factor comparing each hypothesis with the encompassing model increases. Moreover, the hypothesis with a higher number of equality constraints (e.g., Null Hypothesis) will be favoured over hypotheses with a smaller number of equality constraints (e.g., Hierarchy, Independence and Integration Hypothesis).
